Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-09-23 ICLR25/Paper_noProof.tex
valabstraction
update

\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}

@@

In this paper, we studied the LP relaxation of the exact MILP encoding of ReLU-DNNs.
It allowed us to accurately assess when the relaxation of a node impacts very little 
the accuracy, and when it impacts the accuracy importantly, in which case representing it with a binary/integer variable is preferable. This allows us to select a small set of important neurons to encode as a binary variables, while improving accuracy drastically, necessitating $>4$x less integer variables than previous proposals \cite{DivideAndSlide} for the same accuracy. Our empirical studies revealed that Hybrid MILP, can yield highly accurate results, verifying up to 40\% more images than the SOTA ($\alpha,\beta$-Crown, winner of the 4 last VNNComp) within the same runtime. This opens a lot of perspective, among which: verifying efficiently hard instances; certifying robustness of images for larger $\epsilon$ neighbourhoods, verifying global rather than local properties \cite{lipshitz}.
\newpage

>>>>

\newcommand{\val}{{\textrm{value}}}
\newcommand{\Val}{{\textrm{value}}}
\newcommand{\MILP}{{\textrm{MILP}}}
\newcommand{\LP}{{\textrm{LP}}}
\newcommand{\Improve}{\mathrm{Improve}}
\newcommand{\Utility}{\mathrm{Utility}}

@@

In this paper, we studied the LP relaxation of the exact MILP encoding of ReLU-DNNs.
It allowed us to accurately assess when the relaxation of a node impacts very little 
the accuracy, and when it impacts the accuracy importantly, in which case representing it with a binary/integer variable is preferable. This allows us to select a small set of important neurons to encode as a binary variables, while improving accuracy drastically, necessitating $>4$x less integer variables than previous proposals \cite{DivideAndSlide} for the same accuracy. Our empirical studies revealed that Hybrid MILP, can yield highly accurate results, verifying up to 40\% more images than the SOTA ($\alpha,\beta$-Crown, winner of the 4 last VNNComp) within the same runtime. This opens a lot of perspective, among which: verifying efficiently hard instances; certifying robustness of images for larger $\epsilon$ neighbourhoods; verifying global rather than local properties \cite{lipshitz}.

### 2024-09-23 ICLR25/experiments.tex
update

On hard instances (the 6 first DNNs tested), Hybrid MILP is very accurate, only leaving $8\%$-$15\%$ of images undecided, with runtime taking less than $500s$ in average per image, 10 times less even on smaller DNNs. It can scale up to quite large hard CNN instances, such as CNN-B-Adv with 2M parameters and 20K activations.

@@

Last, compared with refined $\beta$-Crown, we can observe three patterns: on the shallowest DNNs (5x100, 5x200), refined $\beta$-Crown can run full MILP on almost all nodes, reaching almost the same accuracy than our Hybrid MILP, but with longer runtime (up to 2x on 5x100). On intermediate DNNs (8x100,8x200), full MILP invoked by refined $\beta$-Crown can only be run on a fraction of the neurons, and the accuracy is not as good as Hybrid MILP, with $6\%-8\%$ more undecided images, while having longer runtime. Last, refined $\beta$-Crown cannot scale to larger instances (6x500, CNN-B-Adv), while Hybrid MILP can.

>>>>

On hard instances (the 6 first DNNs tested), Hybrid MILP is very accurate, only leaving $8\%$-$15\%$ of images undecided, with runtime taking less than $500s$ in average per image, and even 10 times less on smaller DNNs. It can scale up to quite large hard DNNs, such as CNN-B-Adv with 2M parameters and 20K activations.

@@

Last, compared with refined $\beta$-Crown, we can observe three patterns: on the shallowest DNNs (5x100, 5x200), refined $\beta$-Crown can run full MILP on almost all nodes, reaching almost the same accuracy than  Hybrid MILP, but with longer runtime (up to 2x on 5x100). On intermediate DNNs (8x100,8x200), full MILP invoked by refined $\beta$-Crown can only be run on a fraction of the neurons, and the accuracy is not as good as Hybrid MILP, with $6\%-8\%$ more undecided images, while having longer runtime. Last, refined $\beta$-Crown cannot scale to larger instances (6x500, CNN-B-Adv), while Hybrid MILP can.

### 2024-09-23 ICLR25/introduction.tex
update

\begin{table}[b!]
	\centering
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network &  Accuracy & Upper  & Marabou 2.0 & NNenum &  Full MILP  \\ \hline
		MNIST 5$\times$100 & 99\% & 90\% & 28\% & $49\%$ & 40 \%    \\
		$\epsilon = 0.026$ & &  &6200s &  4995s & 6100s
		  \\  \hline
		%  MNIST 5$\times$100 & 99\% & 90\% & 28\% & $46\%$ & 37 \%    \\
		%  TO=2000s & &  & 1250s  & 1194s & 1466s \\  \hline	
	\end{tabular}
\caption{Complete verifiers run on the smallest of the hard DNN with TO = 10 000s. 
%Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and 
Complete verifier barely outperform $\alpha,\beta$-CROWN (40\%, 1026s), despite the much larger runtime, 
still very far from the upper bound.}
\label{table_complete}
\end{table}

The size of the smallest DNN (500 hidden neurons) makes it realistic to be solved by complete verifiers such as Marabou 2.0, NNenum or a Full MILP encoding. 
While they should theoretically be able to close the gap of undecided images,
in practice, Table \ref{table_complete} reveals that even with a very large 10 000s time out per image, none succeds to verify any more image than $\alpha,\beta$-Crown, while the runtime is $>5$ times longer. It seemed pointless to test complete verifier on larger networks.

>>>>

\begin{table}[b!]
	\centering
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network &  Accuracy & Upper  & Marabou 2.0 & NNenum &  Full MILP  \\ \hline
		MNIST 5$\times$100 & 99\% & 90\% & 28\% & $49\%$ & 40 \%    \\
		$\epsilon = 0.026$ & &  &6200s &  4995s & 6100s
		  \\  \hline
		%  MNIST 5$\times$100 & 99\% & 90\% & 28\% & $46\%$ & 37 \%    \\
		%  TO=2000s & &  & 1250s  & 1194s & 1466s \\  \hline	
	\end{tabular}
\caption{Result of complete verifiers on the hard 5x100 with TO = 10 000s. 
%Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and 
Complete verifier barely (9\% out of 50\%) 
outperform $\alpha,\beta$-CROWN (40\%, 1026s), despite much larger runtime.}
\label{table_complete}
\end{table}

The size of the smallest DNN (500 hidden neurons) makes it realistic to be solved by complete verifiers such as Marabou 2.0, NNenum or a Full MILP encoding. 
While they should theoretically be able to close the gap of undecided images,
in practice, Table \ref{table_complete} reveals that only NNenum succeds to verify 9\% more images out of the 50\% undecided images by $\alpha,\beta$-Crown, and with a very large runtime of almost 5000s per image. It seemed pointless to test complete verifiers on larger networks.
